The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2020-06-09 Fsing/FrobeniusThresholdsPaper/response_to_report_C_on_v2.txt

190624-Witt-v2 (FrobeniusThresholds)

The package consists of 3600 lines of code and documentation, so it is
substantial.  It installs and the tests work without error.  The package is in
pretty good shape.

=============================================================================
comments on the code

Please remove the redundant top level directory submission-files/ from the zip
file.

The package TestIdeals has already been published, and there is no need to
include it in the zip file.

Consider choosing a better name for the function "nu", which is simply the
English spelling of a random Greek letter, and thus is not descriptive.
Something with words in it, such as "containmentPower", or even "index".

Consider renaming "fpt" to "pureThreshold" and "isFPT" to "isPureThreshold".
(The "F" for "Frobenius" can be assumed.)

Consider removing redundant parentheses, as in these lines:

		if (I1 == ideal sub(0, S1)) then (

		while ((i < #genList) and (x == sub(0, S1))) do(

		ideal(x) : (ideal(x) : I1)

		annihilator(coker ddual)

	    	(Hom(Hom(M1, S1^1), S1^1))

		if (o.IsGraded==true) then (

If o.IsGraded is always a Boolean value, then instead of

		if (o.IsGraded==true) then (

write

		if o.IsGraded then (

Similarly, instead of

		if (isWellDefined(h) == false) then

you can write

		if not isWellDefined h then

Fix all such comparisons with "true" and "false".

The value of an "if" expression is "null" if the predicate is false, so this
can be omitted:

    else (
        null
    )

Change "if" to "whether" in

       Verifying if special algorithms apply...

The newline characters in verbose messages such as this one

        if o.Verbose then print "\nVerifying if special algorithms apply...";

resulting in double-spaced output, wasting space on the screen and on the pages
of the article.

=============================================================================
comments on the documentation

In

	 "Takes on lists consisting of two numbers"

I don't know what "takes on lists" means.  Maybe say "accepts a list"?

Change

  	"whether to only consider the behavior at the origin"

to

  	"whether to consider only the behavior at the origin"

In

	"toggles verbose output"

you don't mean what you say, since "toggle" replaces false by true and true by
false.  You mean "whether the output is to be verbose"

In

	Consider a Veronese subring (which is étale in codimension 1)

the use of "étale" is incorrect -- it should be applied to ring homomorphism,
not to a ring.

=============================================================================
comments on the article

This

	 the condition that the Frobenius morphism splits as a map of rings

is ambiguous.  Do you mean "is a split surjection" or "is a split injection"?

It would be better if comments in the displayed example code, such as "uses
binary search (default)", were in the surrounding text, because then complete
sentences could be used, and they would be read by the reader in the normal
sequence.  For this purpose, displays could be split, to allow comments in the
text between the two parts.

The displayed example output seems to be a diffferent color, which is
disconcerting.

The displayed example output would look better if it were indented.

The use of "in" in

    	      f is a product of factors in simple normal crossing

sounds peculiar, as does speaking of f instead of its zero set.  Perhaps say
something like

		f defines a simple normal crossing divisor

You use "isSimpleNormalCrossings" in the text, but "isSimpleNormalCrossing" in
the code.

It's adverbs that come after verbs, so change

		can be used independent of

to

		can be used independently of

=============================================================================

### 2020-06-10 Fsing/FrobeniusThresholdsPaper/response_to_report_C_on_v2.txt

The package TestIdeals has already been published, and there is no need to
include it in the zip file.

Consider choosing a better name for the function "nu", which is simply the
English spelling of a random Greek letter, and thus is not descriptive.
Something with words in it, such as "containmentPower", or even "index".

Consider renaming "fpt" to "pureThreshold" and "isFPT" to "isPureThreshold".
(The "F" for "Frobenius" can be assumed.)

*****************************************************************************

Consider removing redundant parentheses...

-----------------------------------------------------------------------------
In progress...
*****************************************************************************

@@

you can write

		if not isWellDefined h then

Fix all such comparisons with "true" and "false".

-----------------------------------------------------------------------------
Done.
*****************************************************************************

@@

The value of an "if" expression is "null" if the predicate is false, so this
can be omitted:

    else (
        null
    )

-----------------------------------------------------------------------------
We'd rather keep the "else null", to make clear the intention to return null.
*****************************************************************************

@@

The newline characters in verbose messages such as this one

        if o.Verbose then print "\nVerifying if special algorithms apply...";

resulting in double-spaced output, wasting space on the screen and on the pages
of the article.

-----------------------------------------------------------------------------
Those were removed.
*****************************************************************************

@@

In

	Consider a Veronese subring (which is étale in codimension 1)

the use of "étale" is incorrect -- it should be applied to ring homomorphism,
not to a ring.

-----------------------------------------------------------------------------
All of the above were fixed.
*****************************************************************************

### 2020-06-10 Fsing/FrobeniusThresholdsPaper/FrobeniusThresholds.tex
Deleted redundant parentheses

\usepackage{fancyvrb}
\DefineVerbatimEnvironment%
{MyVerbatim}{Verbatim}
{formatcom=\color{Violet}}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y];

@@

o6 = 120
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i7 : R = ZZ/5[x,y,z];

@@

o10 = 97
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i11 : nu(5, x^2*y^4 + y^2*z^7 + z^2*x^8, ReturnList => true)

@@

o11 : List
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i12 : R = ZZ/11[x,y,z];

@@

o15 = 1209
\end{MyVerbatim}
}
%\medskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : R = ZZ/3[x,y];

@@

o19 = 26
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : R = ZZ/17[x,y,z];

@@

o23 = 1541642394460
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i24 : R = ZZ/7[x,y];

@@

o27 = 285
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y,z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i5 : R = ZZ/13[x,y];

@@

o8 = false
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

o3 : List
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

o7 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i9 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

@@

o12 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

@@

o15 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

@@

o19 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

o22 : List
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i23 : f = 2*x^10*y^8 + x^4*y^7 - 2*x^3*y^8;

@@

o25 : List
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i26 : time numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
     -- used 0.72874 seconds

@@

o28 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i29 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

o29 : List
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i30 : R = ZZ/7[x,y];

@@

o33 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\subsection{The function \texttt{isSimpleNormalCrossings}} \label{subsec.SNC}
As mentioned earlier, the function \texttt{isSimpleNormalCrossings} verifies whether a polynomial $f$ is a product of factors in simple normal crossing.
Suppose that $f$ has factorization $f_1^{a_i} f_2^{a_2} \cdots f_n^{a_n}$.  Recall that its factors $f_i$ are said to be in
\emph{simple normal crossing} at a point if, locally, they form part of a regular system of parameters.  The function \texttt{isSimpleNormalCrossings} determines whether $f$ has simple normal crossings at the origin by computing the Jacobian matrix of each subset of $\{ f_1, \ldots, f_n \}$ (evaluated at the origin), and checking that these matrices have the expected rank, and that these subsets generate ideals of the appropriate height.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i34 : R = ZZ/7[x,y,z];

@@

o36 = false
\end{MyVerbatim}
}
\smallskip

The function \texttt{isSimpleNormalCrossing} is exposed to the user, so can be used independent of any $F$-pure threshold calculation.
If the user sets its option \texttt{AtOrigin} to \texttt{false} (its default value is \texttt{true}), then the function checks whether the $f_i$ are in simple normal crossing \emph{everywhere}, which can be much slower, since Jacobian ideals are computed.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i37 : R = QQ[x,y,z];

@@

o40 = false
\end{MyVerbatim}
}
\smallskip

>>>>

\usepackage{fancyvrb}
% \DefineVerbatimEnvironment%
% {MyVerbatim}{Verbatim}
% {formatcom=\color{Violet}}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i1 : R = ZZ/11[x,y];

@@

o6 = 120
\end{Verbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i7 : R = ZZ/5[x,y,z];

@@

o10 = 97
\end{Verbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i11 : nu(5, x^2*y^4 + y^2*z^7 + z^2*x^8, ReturnList => true)

@@

o11 : List
\end{Verbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i12 : R = ZZ/11[x,y,z];

@@

o15 = 1209
\end{Verbatim}
}
%\medskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i16 : R = ZZ/3[x,y];

@@

o19 = 26
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i20 : R = ZZ/17[x,y,z];

@@

o23 = 1541642394460
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i24 : R = ZZ/7[x,y];

@@

o27 = 285
\end{Verbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i1 : R = ZZ/11[x,y,z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{Verbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i5 : R = ZZ/13[x,y];

@@

o8 = false
\end{Verbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i1 : R = ZZ/5[x,y,z];

@@

o3 : List
\end{Verbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

o7 : QQ
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i9 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

@@

o12 : QQ
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i13 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

@@

o15 : QQ
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i16 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

@@

o19 : QQ
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i20 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

o22 : List
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i23 : f = 2*x^10*y^8 + x^4*y^7 - 2*x^3*y^8;

@@

o25 : List
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i26 : time numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
     -- used 0.72874 seconds

@@

o28 : QQ
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i29 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

o29 : List
\end{Verbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i30 : R = ZZ/7[x,y];

@@

o33 : QQ
\end{Verbatim}
}
\smallskip

@@

\subsection{The function \texttt{isSimpleNormalCrossing}} \label{subsec.SNC}
As mentioned earlier, the function \texttt{isSimpleNormalCrossing} verifies whether a polynomial $f$ defines a simple normal crossing divisor.
Suppose that $f$ has factorization $f_1^{a_i} f_2^{a_2} \cdots f_n^{a_n}$.  Recall that its factors $f_i$ are said to be in
\emph{simple normal crossing} at a point if, locally, they form part of a regular system of parameters.  The function \texttt{isSimpleNormalCrossing} determines whether $f$ has simple normal crossings at the origin by computing the Jacobian matrix of each subset of $\{ f_1, \ldots, f_n \}$ (evaluated at the origin), and checking that these matrices have the expected rank, and that these subsets generate ideals of the appropriate height.

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i34 : R = ZZ/7[x,y,z];

@@

o36 = false
\end{Verbatim}
}
\smallskip

The function \texttt{isSimpleNormalCrossing} is exposed to the user, so can be used independently of any $F$-pure threshold calculation.
If the user sets its option \texttt{AtOrigin} to \texttt{false} (its default value is \texttt{true}), then the function checks whether the $f_i$ are in simple normal crossing \emph{everywhere}, which can be much slower, since Jacobian ideals are computed.

\smallskip
{\small
\setstretch{.67}
\begin{Verbatim}
i37 : R = QQ[x,y,z];

@@

o40 = false
\end{Verbatim}
}
\smallskip
